
---
title: "Intro DS R Markdown"
author: ""
date: "today"
# date: "`r Sys.Date()`"
output:
  html_document:
    code_folding: hide
    number_sections: false
    toc: yes
    toc_depth: 3
    toc_float: yes
  pdf_document:
    toc: yes
    toc_depth: '3'
---

{r basic, include = F}
# use this function to conveniently load libraries and work smoothly with knitting
# can add quietly=T option to the require() function
# the loadPkg function essentially replaced/substituted two functions install.packages() and library() in one step.
loadPkg = function(x) { if (!require(x,character.only=T, quietly =T)) { install.packages(x,dep=T,repos="http://cran.us.r-project.org"); if(!require(x,character.only=T)) stop("Package not found") } }

# unload/detact package when done using it
unloadPkg = function(pkg, character.only = FALSE) { 
  if(!character.only) { pkg <- as.character(substitute(pkg)) } 
  search_item <- paste("package", pkg,sep = ":") 
  while(search_item %in% search()) { detach(search_item, unload = TRUE, character.only = TRUE) } 
}


```{r setup, include=FALSE}
# some of common options (and the defaults) are: 
# include=T, eval=T, echo=T, results='hide'/'asis'/'markup',..., collapse=F, warning=T, message=T, error=T, cache=T, fig.width=6, fig.height=4, fig.dim=c(6,4) #inches, fig.align='left'/'center','right', 
# knitr::opts_chunk$set(warning = F, results = "markup", message = F)
knitr::opts_chunk$set(warning = F, results = "hide", message = F)
# knitr::opts_chunk$set(include = F)
# knitr::opts_chunk$set(echo = TRUE)
options(scientific=T, digits = 3) 
# options(scipen=9, digits = 3) 
# ‘scipen’: integer. A penalty to be applied when deciding to print numeric values in fixed or exponential notation.  Positive values bias towards fixed and negative towards scientific notation: fixed notation will be preferred unless it is more than ‘scipen’ digits wider.
# use scipen=999 to prevent scientific notation at all times
```
Our team project focuses on the current situation of airlines during the COVID-19 pandemic. We would like to explore, analyze and visualize the current situation with the dataset from the year 2015 to 2020. This enables us to analyze the data before and after COVID. The dataset taken is "US International Air Passengers and Freight Statistics Report",from source kaggle. We have cleaned that dataset by deleted some unnecessary variables and filtering the year from 2015-2020. The resultant dataset contains over 154894 observations. The two important variables in this dataset we have performed the analysis on are the total passengers and total flights. 

### Importing Data

```{r,results='markup'}
library(dplyr)
airlines <- data.frame(read.csv("International_Report_Passengers.csv"))
str(airlines)
```


### Cleaning the data
```{r,results='markup'}

airlines$carriergroup <- as.factor(airlines$carriergroup)
airlines$Year <- as.factor(airlines$Year)
str(airlines)

```


### Dividing the data yearwise
```{r, results='markup'}
s_2015 <- subset(airlines,Year == 2015)

s_2016 <- subset(airlines,Year == 2016)

s_2017 <- subset(airlines,Year == 2017)

s_2018 <- subset(airlines,Year == 2018)

s_2019 <- subset(airlines,Year == 2019)

s_2020 <- subset(airlines,Year == 2020)
```



### SMART question 1: Top 10 busiest airports in the U.S.

```{r,results='markup'}

#Top 10 busy airports 

library(dplyr)

topp_10 <- airlines %>%
	group_by(US_Gateway_AirportCode) %>%
	summarise(Total_Passengers = sum(Total)) %>%
	arrange(desc(Total_Passengers)) %>%
  	top_n(10)
topp_10

```

### SMART question 2: Monthly total number of flights (2015-2020)
```{r,results='markup'}

f_2015 = as.data.frame(table(s_2015$Month))
names(f_2015)[names(f_2015) == "Var1"] <- "Month"
names(f_2015)[names(f_2015) == "Freq"] <- "No_of_Flights"
f_2015


f_2016 = as.data.frame(table(s_2016$Month))
names(f_2016)[names(f_2016) == "Var1"] <- "Month"
names(f_2016)[names(f_2016) == "Freq"] <- "No_of_Flights"
f_2016

f_2017 = as.data.frame(table(s_2017$Month))
names(f_2017)[names(f_2017) == "Var1"] <- "Month"
names(f_2017)[names(f_2017) == "Freq"] <- "No_of_Flights"
f_2017

f_2018 = as.data.frame(table(s_2018$Month))
names(f_2018)[names(f_2018) == "Var1"] <- "Month"
names(f_2018)[names(f_2018) == "Freq"] <- "No_of_Flightss"
f_2018

f_2019 = as.data.frame(table(s_2019$Month))
names(f_2019)[names(f_2019) == "Var1"] <- "Month"
names(f_2019)[names(f_2019) == "Freq"] <- "No_of_Flights"
f_2019

f_2020 = as.data.frame(table(s_2020$Month))
names(f_2020)[names(f_2020) == "Var1"] <- "Month"
names(f_2020)[names(f_2020) == "Freq"] <- "No_of_Flights"
f_2020

```

### Histogram showing monthly flights for each year
```{r,results='markup'}

#hist(s_2015$Month,col = "green", freq = TRUE, main = "Histogram plot showing Monthly Flights for 2015", xlab = "Months",ylab = "Flights",breaks=rep(1:12,each=2)+c(-.4,.4))

#hist(s_2016$Month,col = "green", freq = TRUE, main = "Histogram plot showing Monthly Flights for 2016", xlab = "Months",ylab = "Flights",breaks=rep(1:12,each=2)+c(-.4,.4))

#hist(s_2017$Month,col = "green", freq = TRUE, main = "Histogram plot showing Monthly Flights for 2017", xlab = "Months",ylab = "Flights",breaks=rep(1:12,each=2)+c(-.4,.4))

#hist(s_2018$Month,col = "green", freq = TRUE, main = "Histogram plot showing Monthly Flights for 2018", xlab = "Months",ylab = "Flights",breaks=rep(1:12,each=2)+c(-.4,.4))

#hist(s_2019$Month,col = "green", freq = TRUE, main = "Histogram plot showing Monthly Flights for 2019", xlab = "Months",ylab = "Flights",breaks=rep(1:12,each=2)+c(-.4,.4))

#hist(s_2020$Month,col = "green", freq = TRUE, main = "Histogram plot showing Monthly Flights for 2020", xlab = "Months",ylab = "Flights",breaks=rep(1:12,each=2)+c(-.4,.4))

```

```{r,results='markup'}

install.packages("gridExtra")         
library("gridExtra") 
library(ggplot2)

p1 <- ggplot(s_2015, aes(x=Month)) + geom_histogram(color="black", fill="green", binwidth = 0.5)+ scale_x_continuous(breaks = seq(0, 12, 1))+labs(title = "Histogram plot showing \n Monthly Flights for 2015")+ theme(plot.title = element_text(size = 10, face = "bold"))

p2 <- ggplot(s_2016, aes(x=Month)) + geom_histogram(color="black", fill="green", binwidth = 0.5)+ scale_x_continuous(breaks = seq(0, 12, 1))+labs(title = "Histogram plot showing \n Monthly Flights for 2016")+ theme(plot.title = element_text(size = 10, face = "bold"))

p3 <- ggplot(s_2017, aes(x=Month)) + geom_histogram(color="black", fill="green", binwidth = 0.5)+ scale_x_continuous(breaks = seq(0, 12, 1))+labs(title = "Histogram plot showing \n Monthly Flights for 2017")+ theme(plot.title = element_text(size = 10, face = "bold"))

p4 <- ggplot(s_2018, aes(x=Month)) + geom_histogram(color="black", fill="green", binwidth = 0.5)+ scale_x_continuous(breaks = seq(0, 12, 1))+labs(title = "Histogram plot showing \n Monthly Flights for 2018")+ theme(plot.title = element_text(size = 10, face = "bold"))

p5 <- ggplot(s_2019, aes(x=Month)) + geom_histogram(color="black", fill="green", binwidth = 0.5)+ scale_x_continuous(breaks = seq(0, 12, 1))+labs(title = "Histogram plot showing \n Monthly Flights for 2019")+ theme(plot.title = element_text(size = 10, face = "bold"))

p6 <- ggplot(s_2020, aes(x=Month)) + geom_histogram(color="black", fill="green", binwidth = 0.5)+ scale_x_continuous(breaks = seq(0, 12, 1))+labs(title = "Histogram plot showing \n Monthly Flights for 2020")+ theme(plot.title = element_text(size = 10, face = "bold"))

grid.arrange(p1, p2, p3, p4, p5, p6, ncol = 3, nrow = 2)


```
### Monthly total Number of passengers (2015-2020)
```{r,results='markup'}

p_2015 <- s_2015 %>%
	group_by(Month) %>%
	summarise(Total_Passengers = sum(Total))
p_2015

p_2016 <- s_2016 %>%
	group_by(Month) %>%
	summarise(Total_Passengers = sum(Total))
p_2016

p_2017 <- s_2017 %>%
	group_by(Month) %>%
	summarise(Total_Passengers = sum(Total))
p_2017

p_2018 <- s_2018 %>%
	group_by(Month) %>%
	summarise(Total_Passengers = sum(Total))
p_2018

p_2019 <- s_2019 %>%
	group_by(Month) %>%
	summarise(Total_Passengers = sum(Total))
p_2019

p_2020 <- s_2020 %>%
	group_by(Month) %>%
	summarise(Total_Passengers = sum(Total))
p_2020

```
### Barplots showing Monthly total passengers for each year
```{r,results='markup'}

#barplot(p_2015$Total_Passengers,names.arg=p_2015$Month,xlab="Month",ylab="Total_Passenger",col="blue",main="Monthly passengers for 2015",border="red")


#barplot(p_2016$Total_Passengers,names.arg=p_2016$Month,xlab="Month",ylab="Total_Passenger",col="blue",main="Monthly passengers for 2016",border="red")


#barplot(p_2017$Total_Passengers,names.arg=p_2017$Month,xlab="Month",ylab="Total_Passenger",col="blue",main="Monthly passengers for 2017",border="red")


#barplot(p_2018$Total_Passengers,names.arg=p_2018$Month,xlab="Month",ylab="Total_Passenger",col="blue",main="Monthly passengers for 2018",border="red")


#barplot(p_2019$Total_Passengers,names.arg=p_2019$Month,xlab="Month",ylab="Total_Passenger",col="blue",main="Monthly passengers for 2019",border="red")


#barplot(p_2020$Total_Passengers,names.arg=p_2020$Month,xlab="Month",ylab="Total_Passenger",col="blue",main="Monthly passengers for 2020",border="red")



```


```{r,results='markup'}

# Barplots showing Monthly total passengers for each year
#install.packages("gridExtra")               
library("gridExtra")

p7 <- ggplot(data=p_2015, aes(x=Month, y=Total_Passengers)) + geom_bar(stat="identity", fill="steelblue")+ scale_x_continuous(breaks = seq(0, 12, 1))+labs(title = "Bar Chart showing \n Monthly passengers for 2015")+ theme(plot.title = element_text(size = 10, face = "bold"))

p8 <- ggplot(data=p_2016, aes(x=Month, y=Total_Passengers)) + geom_bar(stat="identity", fill="steelblue")+ scale_x_continuous(breaks = seq(0, 12, 1))+labs(title = "Bar Chart showing \n Monthly passengers for 2016")+ theme(plot.title = element_text(size = 10, face = "bold"))

p9 <- ggplot(data=p_2017, aes(x=Month, y=Total_Passengers)) + geom_bar(stat="identity", fill="steelblue")+ scale_x_continuous(breaks = seq(0, 12, 1))+labs(title = "Bar Chart showing \n Monthly passengers for 2017")+ theme(plot.title = element_text(size = 10, face = "bold"))

p10 <- ggplot(data=p_2018, aes(x=Month, y=Total_Passengers)) + geom_bar(stat="identity", fill="steelblue")+ scale_x_continuous(breaks = seq(0, 12, 1))+labs(title = "Bar Chart showing \n Monthly passengers for 2018")+ theme(plot.title = element_text(size = 10, face = "bold"))

p11 <- ggplot(data=p_2019, aes(x=Month, y=Total_Passengers)) + geom_bar(stat="identity", fill="steelblue")+ scale_x_continuous(breaks = seq(0, 12, 1))+labs(title = "Bar Chart showing \n Monthly passengers for 2019")+ theme(plot.title = element_text(size = 10, face = "bold"))

p12 <- ggplot(data=p_2020, aes(x=Month, y=Total_Passengers)) + geom_bar(stat="identity", fill="steelblue")+ scale_x_continuous(breaks = seq(0, 12, 1))+labs(title = "Bar Chart showing \n Monthly passengers for 2020")+ theme(plot.title = element_text(size = 10, face = "bold"))

grid.arrange(p7, p8, p9, p10, p11, p12, ncol = 3, nrow = 2)
```


### Exploratory Data Analysis
```{r,results='markup'}
# Monthly_Passenger_Data
Monthly_Passenger_Data <- c(p_2015$Total_Passengers, p_2016$Total_Passengers, p_2017$Total_Passengers, p_2018$Total_Passengers, p_2019$Total_Passengers, p_2020$Total_Passengers)
cat("Summary of monthly passenger data from January 2015 to March 2020 \n")
summary(Monthly_Passenger_Data)
cat("\n")

# Monthly_Flight_Data
Monthly_Flight_Data <- c(f_2015$No_of_Flights, f_2016$No_of_Flights, f_2017$No_of_Flights, f_2018$No_of_Flights, f_2019$No_of_Flights, f_2020$No_of_Flights)
cat("Summary of monthly flights data from January 2015 to March 2020 \n")
summary(Monthly_Flight_Data)
cat("\n")

# Standard Deviation
cat("Standard Deviation for monthly passenger data is", sd(Monthly_Passenger_Data))
cat("Standard Deviation for monthly flight data is", sd(Monthly_Flight_Data))

```
```{r,results='markup'}
##EDA :
##Examining the distribution of a categorical variables from the dataset.
##The height of the bar displays how many observations occurred with each x value.

library(tidyverse)
library(ggplot2)
library(dplyr)

ggplot(data = airlines) +
  geom_bar(mapping = aes(x = carriergroup))

##We can also compute these values manually with dplyr::count()

airlines %>%
  count(carriergroup)
```

```{r,results='markup'}  

##Q-Q Plot
##qqnorm(): produces a normal QQ plot of the variable
##qqline(): adds a reference line

qqnorm(Monthly_Flight_Data, pch = 1, frame = TRUE)
qqline(Monthly_Flight_Data, col = "steelblue", lwd = 2)
```

```{r,results='markup'}

##Boxplot 
#boxplot(Monthly_Flight_Data, main="Boxplot for monthly flight data", xlab="No. of flights", ylab="Total", col = "Blue")

Monthly_Flight <- data.frame(Monthly_Flight_Data)
Monthly_Flight %>%
  ggplot(aes(Monthly_Flight_Data)) +
  geom_boxplot(fill="#0000FF") + 
  ggtitle("Boxplot for monthly flight data") +coord_flip()
```
### SMART question 3.1: Impact of COVID-19 on total passengers data

```{r,results='markup'}

month_range = 1:12

plot(month_range, p_2015$Total_Passengers , main="Monthly Variation in number of passengers 2015-2019", type = "b", col = "red", xlab = "Year", ylab = "Total Passengers", ylim=c(13000000,25000000))
par(new = TRUE)
plot(month_range, p_2016$Total_Passengers, type = "b", col = "blue", xlab = "Year", ylab = "Total Passengers", ylim=c(13000000,25000000))
par(new = TRUE)
plot(month_range, p_2017$Total_Passengers, type = "b", col = "green", xlab = "Year", ylab = "Total Passengers", ylim=c(13000000,25000000))
par(new = TRUE)
plot(month_range, p_2018$Total_Passengers, type = "b", col = "black", xlab = "Year", ylab = "Total Passengers", ylim=c(13000000,25000000))
par(new = TRUE)
plot(month_range, p_2019$Total_Passengers, type = "b", col = "magenta", xlab = "Year", ylab = "Total Passengers", ylim=c(13000000,25000000))

axis(1, seq(1,12,1))
legend("topleft",legend=c("Year: 2015", "Year: 2016", "Year: 2017", "Year: 2018", "Year: 2019"), col=c("red", "blue", "green", "black", "magenta"), lty=1:1, cex=0.8)

```
The monthly variation in the number of passengers follow similar trends from 2015 to 2019 which is expected. For example, the peak comes in July for the 4th of July holiday and there is a sharp increase in December as well due to Christmas and the New Year's eve.

```{r,results='markup'}

x_axis = 2015:2020
# Total passengers for the month of January for all the years
y1 <- c(p_2015$Total_Passengers[1], p_2016$Total_Passengers[1], p_2017$Total_Passengers[1], p_2018$Total_Passengers[1], p_2019$Total_Passengers[1], p_2020$Total_Passengers[1])

# Total passengers for the month of February for all the years
y2 <- c(p_2015$Total_Passengers[2], p_2016$Total_Passengers[2], p_2017$Total_Passengers[2], p_2018$Total_Passengers[2], p_2019$Total_Passengers[2], p_2020$Total_Passengers[2])

# Total passengers for the month of March for all the years
y3 <- c(p_2015$Total_Passengers[3], p_2016$Total_Passengers[3], p_2017$Total_Passengers[3], p_2018$Total_Passengers[3], p_2019$Total_Passengers[3], p_2020$Total_Passengers[3])

# Total passengers averaged across all the months for all the years
y4 <- c(mean(p_2015[["Total_Passengers"]]), mean(p_2016[["Total_Passengers"]]), mean(p_2017[["Total_Passengers"]]), mean(p_2018[["Total_Passengers"]]), mean(p_2019[["Total_Passengers"]]), mean(p_2020[["Total_Passengers"]]))
  
plot(x_axis, y1 , main="Yearly Variation in number of passengers", type = "b", col = "red", xlab = "Year", ylab = "Total Passengers", ylim=c(8500000,22000000))
par(new = TRUE)
plot(x_axis, y2 ,type = "b", col = "blue", axes = FALSE, xlab = "Year", ylab = "Total Passengers", ylim=c(8500000,22000000))
par(new = TRUE)
plot(x_axis, y3 ,type = "b", col = "green", axes = FALSE, xlab = "Year", ylab = "Total Passengers", ylim=c(8500000,22000000))
par(new = TRUE)
plot(x_axis, y4 ,type = "b", col = "black", axes = FALSE, xlab = "Year", ylab = "Total Passengers", ylim=c(8500000,22000000))
legend("topleft",legend=c("January", "February", "March", "Average"), col=c("red", "blue", "green", "black"), lty=1:1, cex=0.8)

```

From the plots, we can observe that the yearly variation in total passengers for all the years from 2015 to 2019 follow similar trends. It is followed by a sudden drop in 2020 due to the pandemic. This comparison is done only for months Jan to March as we had data for only those months for 2020. We have also plotted a monthly average for all those years.

### SMART question 3.2: Impact of COVID-19 on total flights data 

```{r,results='markup'}
# Total flights for the month of January for all the years
y5 = c(f_2015$No_of_Flights[1], f_2016$No_of_Flights[1], f_2017$No_of_Flights[1], f_2018$No_of_Flights[1], f_2019$No_of_Flights[1], f_2020$No_of_Flights[1])

# Total flights for the month of February for all the years
y6 = c(f_2015$No_of_Flights[2], f_2016$No_of_Flights[2], f_2017$No_of_Flights[2], f_2018$No_of_Flights[2], f_2019$No_of_Flights[2], f_2020$No_of_Flights[2])

# Total flights for the month of March for all the years
y7 = c(f_2015$No_of_Flights[3], f_2016$No_of_Flights[3], f_2017$No_of_Flights[3], f_2018$No_of_Flights[3], f_2019$No_of_Flights[3], f_2020$No_of_Flights[3])

# Total flights averaged across all the months for all the years
y8 = c(mean(f_2015$No_of_Flights), mean(f_2016$No_of_Flights), mean(f_2017$No_of_Flights), mean(f_2018$No_of_Flights), mean(f_2019$No_of_Flights), mean(f_2020$No_of_Flights))


plot(x_axis, y5 , main="Yearly Variation in number of flights", type = "b", col = "red", xlab = "Year", ylab = "Total Flights", ylim=c(2150,2700))
par(new = TRUE)
plot(x_axis, y6 ,type = "b", col = "blue", axes = FALSE, xlab = "Year", ylab = "Total Flights", ylim=c(2150,2700))
par(new = TRUE)
plot(x_axis, y7 ,type = "b", col = "green", axes = FALSE, xlab = "Year", ylab = "Total Flights", ylim=c(2150,2700))
par(new = TRUE)
plot(x_axis, y8 ,type = "b", col = "black", axes = FALSE, xlab = "Year", ylab = "Total Flights", ylim=c(2150,2700))
legend("topleft",legend=c("January", "February", "March", "Average"), col=c("red", "blue", "green", "black"), lty=1:1, cex=0.8)

```
From the plots, we can observe that the yearly variation in flights for all the years from 2015 to 2019 follow similar trends. It is followed by a sudden drop in 2020. This comparison is done only for months Jan to March as we had data for only those months for 2020. We have also plotted a monthly average for all those years.

### SMART question 4: Post pandemic expectations and predictions

```{r,results='markup'}
#Post Covid


library(dplyr)
library(scales)
library(ggplot2)

fancy_scientific <- function(l) {
     # turn in to character string in scientific notation
     l <- format(l, scientific = TRUE)
     # quote the part before the exponent to keep all the digits
     l <- gsub("^(.*)e", "'\\1'e", l)
     # turn the 'e+' into plotmath format
     l <- gsub("e", "%*%10^", l)
     # return this as an expression
     parse(text=l)
}

year = c(2015, 2016, 2017, 2018, 2019, 2020)

#Average of Monthly passengers by year

Monthly_Mean_By_Year = c(mean(p_2015$Total_Passengers), mean(p_2016$Total_Passengers), mean(p_2017$Total_Passengers), mean(p_2018$Total_Passengers), mean(p_2019$Total_Passengers), mean(p_2020$Total_Passengers))


avg_year<- data.frame(year, Monthly_Mean_By_Year)
avg_year
avg_year %>% 
  ggplot(aes(year, Monthly_Mean_By_Year, colour = year,group=1)) + 
  ggtitle("Plot displaying Monthly Average Passengers from 2015-2020") +
  geom_line()+
  geom_point()+ scale_y_continuous(labels=fancy_scientific) 


#Percent change in the above data

increase15_16 = (mean(p_2016$Total_Passengers)-mean(p_2015$Total_Passengers))/mean(p_2015$Total_Passengers)*100
increase16_17 = (mean(p_2017$Total_Passengers)-mean(p_2016$Total_Passengers))/mean(p_2016$Total_Passengers)*100
increase17_18 = (mean(p_2018$Total_Passengers)-mean(p_2017$Total_Passengers))/mean(p_2017$Total_Passengers)*100
increase18_19 = (mean(p_2019$Total_Passengers)-mean(p_2018$Total_Passengers))/mean(p_2018$Total_Passengers)*100
increase19_20 = (mean(p_2020$Total_Passengers)-mean(p_2019$Total_Passengers))/mean(p_2019$Total_Passengers)*100

percent_increase = c(increase15_16, increase16_17, increase17_18, increase18_19, increase19_20)
year_range = c(2016, 2017, 2018, 2019, 2020 )

per <- data.frame(year_range, percent_increase)
per
per %>%
  ggplot(aes(year_range, percent_increase, colour = year_range,group=1)) + 
  ggtitle("Plot displaying percentage change of passengers 2015-2020") +
  geom_line()+
  geom_point()


#Average of Monthly flights by year

Monthly_mean_flights_year = c(mean(f_2015$No_of_Flights), mean(f_2016$No_of_Flights), mean(f_2017$No_of_Flights), mean(f_2018$No_of_Flights), mean(f_2019$No_of_Flights), mean(f_2020$No_of_Flights) )


avg_year_fl<- data.frame(year, Monthly_mean_flights_year)
avg_year_fl
avg_year %>% 
  ggplot(aes(year, Monthly_mean_flights_year, colour = year,group=1)) + 
  ggtitle("Plot displaying Monthly Average flights from 2015-2020") +
  geom_line()+
  geom_point()+ scale_y_continuous(labels=fancy_scientific) 

```
### Predicting number of flights required to accommodate higher number of passengers post COVID

Flights and passengers data is grouped according to the airport.

```{r,results='markup'}
#Airport wise totals
library(dplyr)
Airport_totals <- airlines %>%
	group_by(US_Gateway_AirportCode) %>%
	summarise(Total_Pass = sum(Total), Tot_flights = n())

Airport_totals
# Correlation between number of departures and number of enplaned passengers 
cor.test(Airport_totals$Total_Pass, Airport_totals$Tot_flights)

```

After performing Pearson's correlation test, we found out that the correlation coefficient comes out to be 0.966 which means this is a strong positive relation between total passengers and the total number of departures. An increase in one results in an increase in the later.


```{r,results='markup'}
# Scatterplot
plot(Airport_totals$Total_Pass, Airport_totals$Tot_flights, main="Departures vs Passengers", xlab="Total passengers ", ylab="Total departures", pch=19)
# Best fit line
abline(lm(Airport_totals$Tot_flights~Airport_totals$Total_Pass), col="red")

#linear regression
flight_pass_regression = lm(Tot_flights~Total_Pass, data = Airport_totals)
summary(flight_pass_regression)

```

The R-squared value of the linear regression model is 0.932 which means the model can explain 93.2% of the total variability in the data. Also the best fit line shown in the scatterplot complement this result. The p-value for Total passengers is 2*e-16. A very small value like this means that the variables we chose are excellent to create a meaningful model.

With the help of the linear regression model, we can predict total number of flights required to accommodate increased number of passengers post COVID. Following is the equation for the same:

Tot_flights = 82.0 + (9.28e-05)*Total_Pass

### Conclusion

-In this project, we analyzed a dataset containing daily flights and passenger’s data for various airlines. Our analysis started with us combining the dates into their respective months and dividing the data year wise from 2015 to 2020. 

-Next step was to form separate databases for monthly total of passengers and flights and running analyses on them. Our first tangible and publishable result was finding top ten busiest airports. 

-We continued our analysis by plotting histograms, bar plots, Q-Q plots, and boxplots to determine the distribution of the passengers and flights data. These plots were studied, and we found out that the monthly variations for both the variables followed a similar trend (rise and fall in the values) irrespective of the year, till 2019. This was to be expected as the holidays fall on the same dates every year. 

-After plotting the values for the year 2020, we found out a sudden decrease in the values in the month of March, which was expected as well, due to the COVID-19 outbreak.

-We plotted the percentage change in the values of passengers and flights and arrived on the conclusion that we can expect the percentages to be in the same range post COVID. 

-We formed a linear regression model between total enplaned passengers and total number of departures after performing a correlation test on them. 

-Our final task was to predict the total number of flights an airport would need, to accommodate an increased number of passengers. This prediction would greatly help the airports to manage and allocate their resources post COVID. The regression model we formed helped us do the same. 

-Thus, we finished our analysis of the dataset with useful conclusions and results.
