
---
title: "Intro DS R Markdown"
author: ""
date: "today"
# date: "`r Sys.Date()`"
output:
  html_document:
    code_folding: hide
    number_sections: false
    toc: yes
    toc_depth: 3
    toc_float: yes
  pdf_document:
    toc: yes
    toc_depth: '3'
---

{r basic, include = F}
# use this function to conveniently load libraries and work smoothly with knitting
# can add quietly=T option to the require() function
# the loadPkg function essentially replaced/substituted two functions install.packages() and library() in one step.
loadPkg = function(x) { if (!require(x,character.only=T, quietly =T)) { install.packages(x,dep=T,repos="http://cran.us.r-project.org"); if(!require(x,character.only=T)) stop("Package not found") } }

# unload/detact package when done using it
unloadPkg = function(pkg, character.only = FALSE) { 
  if(!character.only) { pkg <- as.character(substitute(pkg)) } 
  search_item <- paste("package", pkg,sep = ":") 
  while(search_item %in% search()) { detach(search_item, unload = TRUE, character.only = TRUE) } 
}


{r setup, include=FALSE}
# some of common options (and the defaults) are: 
# include=T, eval=T, echo=T, results='hide'/'asis'/'markup',..., collapse=F, warning=T, message=T, error=T, cache=T, fig.width=6, fig.height=4, fig.dim=c(6,4) #inches, fig.align='left'/'center','right', 
# knitr::opts_chunk$set(warning = F, results = "markup", message = F)
knitr::opts_chunk$set(warning = F, results = "hide", message = F)
# knitr::opts_chunk$set(include = F)
# knitr::opts_chunk$set(echo = TRUE)
options(scientific=T, digits = 3) 
# options(scipen=9, digits = 3) 
# ‘scipen’: integer. A penalty to be applied when deciding to print numeric values in fixed or exponential notation.  Positive values bias towards fixed and negative towards scientific notation: fixed notation will be preferred unless it is more than ‘scipen’ digits wider.
# use scipen=999 to prevent scientific notation at all times
```{r,results='markup'}
library(dplyr)
airlines <- data.frame(read.csv("International_Report_Passengers.csv"))
str(airlines)
```



```{r,results='markup'}

airlines$carriergroup <- as.factor(airlines$carriergroup)
airlines$Year <- as.factor(airlines$Year)
str(airlines)

```


### Dividing the data yearwise
```{r, results='markup'}
s_2015 <- subset(airlines,Year == 2015)

s_2016 <- subset(airlines,Year == 2016)

s_2017 <- subset(airlines,Year == 2017)

s_2018 <- subset(airlines,Year == 2018)

s_2019 <- subset(airlines,Year == 2019)

s_2020 <- subset(airlines,Year == 2020)
```



### Top 10 busiest airports in the U.S.

```{r,results='markup'}

#Top 10 busy airports 

library(dplyr)

topp_10 <- airlines %>%
	group_by(US_Gateway_AirportCode) %>%
	summarise(Total_Passengers = sum(Total)) %>%
	arrange(desc(Total_Passengers)) %>%
  	top_n(10)
topp_10

```

### Monthly total number of flights.(2015-2020)
```{r,results='markup'}

f_2015 = as.data.frame(table(s_2015$Month))
names(f_2015)[names(f_2015) == "Var1"] <- "Month"
names(f_2015)[names(f_2015) == "Freq"] <- "No_of_Flights"
f_2015


f_2016 = as.data.frame(table(s_2016$Month))
names(f_2016)[names(f_2016) == "Var1"] <- "Month"
names(f_2016)[names(f_2016) == "Freq"] <- "No_of_Flights"
f_2016

f_2017 = as.data.frame(table(s_2017$Month))
names(f_2017)[names(f_2017) == "Var1"] <- "Month"
names(f_2017)[names(f_2017) == "Freq"] <- "No_of_Flights"
f_2017

f_2018 = as.data.frame(table(s_2018$Month))
names(f_2018)[names(f_2018) == "Var1"] <- "Month"
names(f_2018)[names(f_2018) == "Freq"] <- "No_of_Flightss"
f_2018

f_2019 = as.data.frame(table(s_2019$Month))
names(f_2019)[names(f_2019) == "Var1"] <- "Month"
names(f_2019)[names(f_2019) == "Freq"] <- "No_of_Flights"
f_2019

f_2020 = as.data.frame(table(s_2020$Month))
names(f_2020)[names(f_2020) == "Var1"] <- "Month"
names(f_2020)[names(f_2020) == "Freq"] <- "No_of_Flights"
f_2020

```


```{r,results='markup'}

# Histogram showing monthly flights for each year

#hist(s_2015$Month,col = "green", freq = TRUE, main = "Histogram plot showing Monthly Flights for 2015", xlab = "Months",ylab = "Flights",breaks=rep(1:12,each=2)+c(-.4,.4))

#hist(s_2016$Month,col = "green", freq = TRUE, main = "Histogram plot showing Monthly Flights for 2016", xlab = "Months",ylab = "Flights",breaks=rep(1:12,each=2)+c(-.4,.4))

#hist(s_2017$Month,col = "green", freq = TRUE, main = "Histogram plot showing Monthly Flights for 2017", xlab = "Months",ylab = "Flights",breaks=rep(1:12,each=2)+c(-.4,.4))

#hist(s_2018$Month,col = "green", freq = TRUE, main = "Histogram plot showing Monthly Flights for 2018", xlab = "Months",ylab = "Flights",breaks=rep(1:12,each=2)+c(-.4,.4))

#hist(s_2019$Month,col = "green", freq = TRUE, main = "Histogram plot showing Monthly Flights for 2019", xlab = "Months",ylab = "Flights",breaks=rep(1:12,each=2)+c(-.4,.4))

#hist(s_2020$Month,col = "green", freq = TRUE, main = "Histogram plot showing Monthly Flights for 2020", xlab = "Months",ylab = "Flights",breaks=rep(1:12,each=2)+c(-.4,.4))

```

```{r,results='markup'}

#install.packages("gridExtra")         
library("gridExtra") 
library(ggplot2)

p1 <- ggplot(s_2015, aes(x=Month)) + geom_histogram(color="black", fill="green", binwidth = 0.5)+ scale_x_continuous(breaks = seq(0, 12, 1))+labs(title = "Histogram plot showing \n Monthly Flights for 2015")+ theme(plot.title = element_text(size = 10, face = "bold"))

p2 <- ggplot(s_2016, aes(x=Month)) + geom_histogram(color="black", fill="green", binwidth = 0.5)+ scale_x_continuous(breaks = seq(0, 12, 1))+labs(title = "Histogram plot showing \n Monthly Flights for 2016")+ theme(plot.title = element_text(size = 10, face = "bold"))

p3 <- ggplot(s_2017, aes(x=Month)) + geom_histogram(color="black", fill="green", binwidth = 0.5)+ scale_x_continuous(breaks = seq(0, 12, 1))+labs(title = "Histogram plot showing \n Monthly Flights for 2017")+ theme(plot.title = element_text(size = 10, face = "bold"))

p4 <- ggplot(s_2018, aes(x=Month)) + geom_histogram(color="black", fill="green", binwidth = 0.5)+ scale_x_continuous(breaks = seq(0, 12, 1))+labs(title = "Histogram plot showing \n Monthly Flights for 2018")+ theme(plot.title = element_text(size = 10, face = "bold"))

p5 <- ggplot(s_2019, aes(x=Month)) + geom_histogram(color="black", fill="green", binwidth = 0.5)+ scale_x_continuous(breaks = seq(0, 12, 1))+labs(title = "Histogram plot showing \n Monthly Flights for 2019")+ theme(plot.title = element_text(size = 10, face = "bold"))

p6 <- ggplot(s_2020, aes(x=Month)) + geom_histogram(color="black", fill="green", binwidth = 0.5)+ scale_x_continuous(breaks = seq(0, 12, 1))+labs(title = "Histogram plot showing \n Monthly Flights for 2020")+ theme(plot.title = element_text(size = 10, face = "bold"))

grid.arrange(p1, p2, p3, p4, p5, p6, ncol = 3, nrow = 2)


```

```{r,results='markup'}

#Number of Passengers Monthly

p_2015 <- s_2015 %>%
	group_by(Month) %>%
	summarise(Total_Passengers = sum(Total))
p_2015

p_2016 <- s_2016 %>%
	group_by(Month) %>%
	summarise(Total_Passengers = sum(Total))
p_2016

p_2017 <- s_2017 %>%
	group_by(Month) %>%
	summarise(Total_Passengers = sum(Total))
p_2017

p_2018 <- s_2018 %>%
	group_by(Month) %>%
	summarise(Total_Passengers = sum(Total))
p_2018

p_2019 <- s_2019 %>%
	group_by(Month) %>%
	summarise(Total_Passengers = sum(Total))
p_2019

p_2020 <- s_2020 %>%
	group_by(Month) %>%
	summarise(Total_Passengers = sum(Total))
p_2020

```

```{r,results='markup'}

# Barplots showing Monthly total passengers for each year

#barplot(p_2015$Total_Passengers,names.arg=p_2015$Month,xlab="Month",ylab="Total_Passenger",col="blue",main="Monthly passengers for 2015",border="red")


#barplot(p_2016$Total_Passengers,names.arg=p_2016$Month,xlab="Month",ylab="Total_Passenger",col="blue",main="Monthly passengers for 2016",border="red")


#barplot(p_2017$Total_Passengers,names.arg=p_2017$Month,xlab="Month",ylab="Total_Passenger",col="blue",main="Monthly passengers for 2017",border="red")


#barplot(p_2018$Total_Passengers,names.arg=p_2018$Month,xlab="Month",ylab="Total_Passenger",col="blue",main="Monthly passengers for 2018",border="red")


#barplot(p_2019$Total_Passengers,names.arg=p_2019$Month,xlab="Month",ylab="Total_Passenger",col="blue",main="Monthly passengers for 2019",border="red")


#barplot(p_2020$Total_Passengers,names.arg=p_2020$Month,xlab="Month",ylab="Total_Passenger",col="blue",main="Monthly passengers for 2020",border="red")



```


```{r,results='markup'}

# Barplots showing Monthly total passengers for each year
install.packages("gridExtra")               # Install gridExtra package
library("gridExtra")

p7 <- ggplot(data=p_2015, aes(x=Month, y=Total_Passengers)) + geom_bar(stat="identity", fill="steelblue")+ scale_x_continuous(breaks = seq(0, 12, 1))+labs(title = "Histogram plot showing \n Monthly Flights for 2015")+ theme(plot.title = element_text(size = 10, face = "bold"))

p8 <- ggplot(data=p_2016, aes(x=Month, y=Total_Passengers)) + geom_bar(stat="identity", fill="steelblue")+ scale_x_continuous(breaks = seq(0, 12, 1))+labs(title = "Histogram plot showing \n Monthly Flights for 2016")+ theme(plot.title = element_text(size = 10, face = "bold"))

p9 <- ggplot(data=p_2017, aes(x=Month, y=Total_Passengers)) + geom_bar(stat="identity", fill="steelblue")+ scale_x_continuous(breaks = seq(0, 12, 1))+labs(title = "Histogram plot showing \n Monthly Flights for 2017")+ theme(plot.title = element_text(size = 10, face = "bold"))

p10 <- ggplot(data=p_2018, aes(x=Month, y=Total_Passengers)) + geom_bar(stat="identity", fill="steelblue")+ scale_x_continuous(breaks = seq(0, 12, 1))+labs(title = "Histogram plot showing \n Monthly Flights for 2018")+ theme(plot.title = element_text(size = 10, face = "bold"))

p11 <- ggplot(data=p_2019, aes(x=Month, y=Total_Passengers)) + geom_bar(stat="identity", fill="steelblue")+ scale_x_continuous(breaks = seq(0, 12, 1))+labs(title = "Histogram plot showing \n Monthly Flights for 2019")+ theme(plot.title = element_text(size = 10, face = "bold"))

p12 <- ggplot(data=p_2020, aes(x=Month, y=Total_Passengers)) + geom_bar(stat="identity", fill="steelblue")+ scale_x_continuous(breaks = seq(0, 12, 1))+labs(title = "Histogram plot showing \n Monthly Flights for 2020")+ theme(plot.title = element_text(size = 10, face = "bold"))

grid.arrange(p7, p8, p9, p10, p11, p12, ncol = 3, nrow = 2)
```
```{r,results='markup'}
Monthly_Passenger_Data <- c(p_2015$Total_Passengers, p_2016$Total_Passengers, p_2017$Total_Passengers, p_2018$Total_Passengers, p_2019$Total_Passengers, p_2020$Total_Passengers)
summary(Monthly_Passenger_Data)
cat("\n")

Monthly_Flight_Data <- c(f_2015$No_of_Flights, f_2016$No_of_Flights, f_2017$No_of_Flights, f_2018$No_of_Flights, f_2019$No_of_Flights, f_2020$No_of_Flights)
# Monthly_Flight_Data
summary(Monthly_Flight_Data)
cat("\n")

# Standard Deviation
sd(Monthly_Passenger_Data)
sd(Monthly_Flight_Data)

```
