
---
title: "Intro DS R Markdown"
author: ""
date: "today"
# date: "`r Sys.Date()`"
output:
  html_document:
    code_folding: hide
    number_sections: false
    toc: yes
    toc_depth: 3
    toc_float: yes
  pdf_document:
    toc: yes
    toc_depth: '3'
---

{r basic, include = F}
# use this function to conveniently load libraries and work smoothly with knitting
# can add quietly=T option to the require() function
# the loadPkg function essentially replaced/substituted two functions install.packages() and library() in one step.
loadPkg = function(x) { if (!require(x,character.only=T, quietly =T)) { install.packages(x,dep=T,repos="http://cran.us.r-project.org"); if(!require(x,character.only=T)) stop("Package not found") } }

# unload/detact package when done using it
unloadPkg = function(pkg, character.only = FALSE) { 
  if(!character.only) { pkg <- as.character(substitute(pkg)) } 
  search_item <- paste("package", pkg,sep = ":") 
  while(search_item %in% search()) { detach(search_item, unload = TRUE, character.only = TRUE) } 
}


{r setup, include=FALSE}
# some of common options (and the defaults) are: 
# include=T, eval=T, echo=T, results='hide'/'asis'/'markup',..., collapse=F, warning=T, message=T, error=T, cache=T, fig.width=6, fig.height=4, fig.dim=c(6,4) #inches, fig.align='left'/'center','right', 
# knitr::opts_chunk$set(warning = F, results = "markup", message = F)
knitr::opts_chunk$set(warning = F, results = "hide", message = F)
# knitr::opts_chunk$set(include = F)
# knitr::opts_chunk$set(echo = TRUE)
options(scientific=T, digits = 3) 
# options(scipen=9, digits = 3) 
# ‘scipen’: integer. A penalty to be applied when deciding to print numeric values in fixed or exponential notation.  Positive values bias towards fixed and negative towards scientific notation: fixed notation will be preferred unless it is more than ‘scipen’ digits wider.
# use scipen=999 to prevent scientific notation at all times
```{r,results='markup'}
library(dplyr)
airlines <- data.frame(read.csv("International_Report_Passengers.csv"))
str(airlines)
```



```{r,results='markup'}

airlines$carriergroup <- as.factor(airlines$carriergroup)
airlines$Year <- as.factor(airlines$Year)
str(airlines)

```


### Dividing the data yearwise
```{r, results='markup'}
s_2015 <- subset(airlines,Year == 2015)

s_2016 <- subset(airlines,Year == 2016)

s_2017 <- subset(airlines,Year == 2017)

s_2018 <- subset(airlines,Year == 2018)

s_2019 <- subset(airlines,Year == 2019)

s_2020 <- subset(airlines,Year == 2020)
```



### Top 10 busiest airports in the U.S.

```{r,results='markup'}

#Top 10 busy airports 

library(dplyr)

topp_10 <- airlines %>%
	group_by(US_Gateway_AirportCode) %>%
	summarise(Total_Passengers = sum(Total)) %>%
	arrange(desc(Total_Passengers)) %>%
  	top_n(10)
topp_10

```

### Monthly total number of flights.(2015-2020)
```{r,results='markup'}

f_2015 = as.data.frame(table(s_2015$Month))
names(f_2015)[names(f_2015) == "Var1"] <- "Month"
names(f_2015)[names(f_2015) == "Freq"] <- "No_of_Flights"
f_2015


f_2016 = as.data.frame(table(s_2016$Month))
names(f_2016)[names(f_2016) == "Var1"] <- "Month"
names(f_2016)[names(f_2016) == "Freq"] <- "No_of_Flights"
f_2016

f_2017 = as.data.frame(table(s_2017$Month))
names(f_2017)[names(f_2017) == "Var1"] <- "Month"
names(f_2017)[names(f_2017) == "Freq"] <- "No_of_Flights"
f_2017

f_2018 = as.data.frame(table(s_2018$Month))
names(f_2018)[names(f_2018) == "Var1"] <- "Month"
names(f_2018)[names(f_2018) == "Freq"] <- "No_of_Flightss"
f_2018

f_2019 = as.data.frame(table(s_2019$Month))
names(f_2019)[names(f_2019) == "Var1"] <- "Month"
names(f_2019)[names(f_2019) == "Freq"] <- "No_of_Flights"
f_2019

f_2020 = as.data.frame(table(s_2020$Month))
names(f_2020)[names(f_2020) == "Var1"] <- "Month"
names(f_2020)[names(f_2020) == "Freq"] <- "No_of_Flights"
f_2020

```


```{r,results='markup'}

# Histogram showing monthly flights for each year

#hist(s_2015$Month,col = "green", freq = TRUE, main = "Histogram plot showing Monthly Flights for 2015", xlab = "Months",ylab = "Flights",breaks=rep(1:12,each=2)+c(-.4,.4))

#hist(s_2016$Month,col = "green", freq = TRUE, main = "Histogram plot showing Monthly Flights for 2016", xlab = "Months",ylab = "Flights",breaks=rep(1:12,each=2)+c(-.4,.4))

#hist(s_2017$Month,col = "green", freq = TRUE, main = "Histogram plot showing Monthly Flights for 2017", xlab = "Months",ylab = "Flights",breaks=rep(1:12,each=2)+c(-.4,.4))

#hist(s_2018$Month,col = "green", freq = TRUE, main = "Histogram plot showing Monthly Flights for 2018", xlab = "Months",ylab = "Flights",breaks=rep(1:12,each=2)+c(-.4,.4))

#hist(s_2019$Month,col = "green", freq = TRUE, main = "Histogram plot showing Monthly Flights for 2019", xlab = "Months",ylab = "Flights",breaks=rep(1:12,each=2)+c(-.4,.4))

#hist(s_2020$Month,col = "green", freq = TRUE, main = "Histogram plot showing Monthly Flights for 2020", xlab = "Months",ylab = "Flights",breaks=rep(1:12,each=2)+c(-.4,.4))

```

```{r,results='markup'}

install.packages("gridExtra")         
library("gridExtra") 
library(ggplot2)

p1 <- ggplot(s_2015, aes(x=Month)) + geom_histogram(color="black", fill="green", binwidth = 0.5)+ scale_x_continuous(breaks = seq(0, 12, 1))+labs(title = "Histogram plot showing \n Monthly Flights for 2015")+ theme(plot.title = element_text(size = 10, face = "bold"))

p2 <- ggplot(s_2016, aes(x=Month)) + geom_histogram(color="black", fill="green", binwidth = 0.5)+ scale_x_continuous(breaks = seq(0, 12, 1))+labs(title = "Histogram plot showing \n Monthly Flights for 2016")+ theme(plot.title = element_text(size = 10, face = "bold"))

p3 <- ggplot(s_2017, aes(x=Month)) + geom_histogram(color="black", fill="green", binwidth = 0.5)+ scale_x_continuous(breaks = seq(0, 12, 1))+labs(title = "Histogram plot showing \n Monthly Flights for 2017")+ theme(plot.title = element_text(size = 10, face = "bold"))

p4 <- ggplot(s_2018, aes(x=Month)) + geom_histogram(color="black", fill="green", binwidth = 0.5)+ scale_x_continuous(breaks = seq(0, 12, 1))+labs(title = "Histogram plot showing \n Monthly Flights for 2018")+ theme(plot.title = element_text(size = 10, face = "bold"))

p5 <- ggplot(s_2019, aes(x=Month)) + geom_histogram(color="black", fill="green", binwidth = 0.5)+ scale_x_continuous(breaks = seq(0, 12, 1))+labs(title = "Histogram plot showing \n Monthly Flights for 2019")+ theme(plot.title = element_text(size = 10, face = "bold"))

p6 <- ggplot(s_2020, aes(x=Month)) + geom_histogram(color="black", fill="green", binwidth = 0.5)+ scale_x_continuous(breaks = seq(0, 12, 1))+labs(title = "Histogram plot showing \n Monthly Flights for 2020")+ theme(plot.title = element_text(size = 10, face = "bold"))

grid.arrange(p1, p2, p3, p4, p5, p6, ncol = 3, nrow = 2)


```

```{r,results='markup'}

#Number of Passengers Monthly

p_2015 <- s_2015 %>%
	group_by(Month) %>%
	summarise(Total_Passengers = sum(Total))
p_2015

p_2016 <- s_2016 %>%
	group_by(Month) %>%
	summarise(Total_Passengers = sum(Total))
p_2016

p_2017 <- s_2017 %>%
	group_by(Month) %>%
	summarise(Total_Passengers = sum(Total))
p_2017

p_2018 <- s_2018 %>%
	group_by(Month) %>%
	summarise(Total_Passengers = sum(Total))
p_2018

p_2019 <- s_2019 %>%
	group_by(Month) %>%
	summarise(Total_Passengers = sum(Total))
p_2019

p_2020 <- s_2020 %>%
	group_by(Month) %>%
	summarise(Total_Passengers = sum(Total))
p_2020

```

```{r,results='markup'}

# Barplots showing Monthly total passengers for each year

#barplot(p_2015$Total_Passengers,names.arg=p_2015$Month,xlab="Month",ylab="Total_Passenger",col="blue",main="Monthly passengers for 2015",border="red")


#barplot(p_2016$Total_Passengers,names.arg=p_2016$Month,xlab="Month",ylab="Total_Passenger",col="blue",main="Monthly passengers for 2016",border="red")


#barplot(p_2017$Total_Passengers,names.arg=p_2017$Month,xlab="Month",ylab="Total_Passenger",col="blue",main="Monthly passengers for 2017",border="red")


#barplot(p_2018$Total_Passengers,names.arg=p_2018$Month,xlab="Month",ylab="Total_Passenger",col="blue",main="Monthly passengers for 2018",border="red")


#barplot(p_2019$Total_Passengers,names.arg=p_2019$Month,xlab="Month",ylab="Total_Passenger",col="blue",main="Monthly passengers for 2019",border="red")


#barplot(p_2020$Total_Passengers,names.arg=p_2020$Month,xlab="Month",ylab="Total_Passenger",col="blue",main="Monthly passengers for 2020",border="red")



```


```{r,results='markup'}

# Barplots showing Monthly total passengers for each year
#install.packages("gridExtra")               
library("gridExtra")

p7 <- ggplot(data=p_2015, aes(x=Month, y=Total_Passengers)) + geom_bar(stat="identity", fill="steelblue")+ scale_x_continuous(breaks = seq(0, 12, 1))+labs(title = "Bar Chart showing \n Monthly passengers for 2015")+ theme(plot.title = element_text(size = 10, face = "bold"))

p8 <- ggplot(data=p_2016, aes(x=Month, y=Total_Passengers)) + geom_bar(stat="identity", fill="steelblue")+ scale_x_continuous(breaks = seq(0, 12, 1))+labs(title = "Bar Chart showing \n Monthly passengers for 2016")+ theme(plot.title = element_text(size = 10, face = "bold"))

p9 <- ggplot(data=p_2017, aes(x=Month, y=Total_Passengers)) + geom_bar(stat="identity", fill="steelblue")+ scale_x_continuous(breaks = seq(0, 12, 1))+labs(title = "Bar Chart showing \n Monthly passengers for 2017")+ theme(plot.title = element_text(size = 10, face = "bold"))

p10 <- ggplot(data=p_2018, aes(x=Month, y=Total_Passengers)) + geom_bar(stat="identity", fill="steelblue")+ scale_x_continuous(breaks = seq(0, 12, 1))+labs(title = "Bar Chart showing \n Monthly passengers for 2018")+ theme(plot.title = element_text(size = 10, face = "bold"))

p11 <- ggplot(data=p_2019, aes(x=Month, y=Total_Passengers)) + geom_bar(stat="identity", fill="steelblue")+ scale_x_continuous(breaks = seq(0, 12, 1))+labs(title = "Bar Chart showing \n Monthly passengers for 2019")+ theme(plot.title = element_text(size = 10, face = "bold"))

p12 <- ggplot(data=p_2020, aes(x=Month, y=Total_Passengers)) + geom_bar(stat="identity", fill="steelblue")+ scale_x_continuous(breaks = seq(0, 12, 1))+labs(title = "Bar Chart showing \n Monthly passengers for 2020")+ theme(plot.title = element_text(size = 10, face = "bold"))

grid.arrange(p7, p8, p9, p10, p11, p12, ncol = 3, nrow = 2)
```


### Exploratory Data Analysis
```{r,results='markup'}
# Monthly_Passenger_Data
Monthly_Passenger_Data <- c(p_2015$Total_Passengers, p_2016$Total_Passengers, p_2017$Total_Passengers, p_2018$Total_Passengers, p_2019$Total_Passengers, p_2020$Total_Passengers)
cat("Summary of monthly passenger data from January 2015 to March 2020 \n")
summary(Monthly_Passenger_Data)
cat("\n")

# Monthly_Flight_Data
Monthly_Flight_Data <- c(f_2015$No_of_Flights, f_2016$No_of_Flights, f_2017$No_of_Flights, f_2018$No_of_Flights, f_2019$No_of_Flights, f_2020$No_of_Flights)
cat("Summary of monthly flights data from January 2015 to March 2020 \n")
summary(Monthly_Flight_Data)
cat("\n")

# Standard Deviation
cat("Standard Deviation for monthly passenger data is", sd(Monthly_Passenger_Data))
cat("Standard Deviation for monthly flight data is", sd(Monthly_Flight_Data))

```
```{r,results='markup'}
##EDA :

##Examining the distribution of a categorical variables from the dataset.
##The height of the bar displays how many observations occurred with each x value.

library(tidyverse)
library(ggplot2)
library(dplyr)

ggplot(data = airlines) +
  geom_bar(mapping = aes(x = carriergroup))

##We can also compute these values manually with dplyr::count()

airlines %>%
  count(carriergroup)
  
  ##Q-Q Plot
##qqnorm(): produces a normal QQ plot of the variable
##qqline(): adds a reference line

qqnorm(Monthly_Flight_Data, pch = 1, frame = TRUE)
qqline(Monthly_Flight_Data, col = "steelblue", lwd = 2)

##Boxplot 
boxplot(Monthly_Flight_Data, main="Boxplot for monthly flight data", xlab="No. of flights", ylab="Total", col = "Blue")
```
```{r,results='markup'}
# Impact of COVID-19 on total passengers data

x_axis = 2015:2020

# Total passengers for the month of January for all the years
y1 <- c(p_2015$Total_Passengers[1], p_2016$Total_Passengers[1], p_2017$Total_Passengers[1], p_2018$Total_Passengers[1], p_2019$Total_Passengers[1], p_2020$Total_Passengers[1])

# Total passengers for the month of February for all the years
y2 <- c(p_2015$Total_Passengers[2], p_2016$Total_Passengers[2], p_2017$Total_Passengers[2], p_2018$Total_Passengers[2], p_2019$Total_Passengers[2], p_2020$Total_Passengers[2])

# Total passengers for the month of March for all the years
y3 <- c(p_2015$Total_Passengers[3], p_2016$Total_Passengers[3], p_2017$Total_Passengers[3], p_2018$Total_Passengers[3], p_2019$Total_Passengers[3], p_2020$Total_Passengers[3])

# Total passengers averaged across all the months for all the years
y4 <- c(mean(p_2015[["Total_Passengers"]]), mean(p_2016[["Total_Passengers"]]), mean(p_2017[["Total_Passengers"]]), mean(p_2018[["Total_Passengers"]]), mean(p_2019[["Total_Passengers"]]), mean(p_2020[["Total_Passengers"]]))
  
plot(x_axis, y1 , main="Yearly Variation in number of passengers", type = "b", col = "red", xlab = "Year", ylab = "Total Passengers", ylim=c(8500000,22000000))
par(new = TRUE)
plot(x_axis, y2 ,type = "b", col = "blue", axes = FALSE, xlab = "Year", ylab = "Total Passengers", ylim=c(8500000,22000000))
par(new = TRUE)
plot(x_axis, y3 ,type = "b", col = "green", axes = FALSE, xlab = "Year", ylab = "Total Passengers", ylim=c(8500000,22000000))
par(new = TRUE)
plot(x_axis, y4 ,type = "b", col = "black", axes = FALSE, xlab = "Year", ylab = "Total Passengers", ylim=c(8500000,22000000))
legend("topleft",legend=c("January", "February", "March", "Average"), col=c("red", "blue", "green", "black"), lty=1:1, cex=0.8)



```{r}

# Impact of COVID-19 on total flights data 

# Total flights for the month of January for all the years
y5 = c(f_2015$No_of_Flights[1], f_2016$No_of_Flights[1], f_2017$No_of_Flights[1], f_2018$No_of_Flights[1], f_2019$No_of_Flights[1], f_2020$No_of_Flights[1])

# Total flights for the month of February for all the years
y6 = c(f_2015$No_of_Flights[2], f_2016$No_of_Flights[2], f_2017$No_of_Flights[2], f_2018$No_of_Flights[2], f_2019$No_of_Flights[2], f_2020$No_of_Flights[2])

# Total flights for the month of March for all the years
y7 = c(f_2015$No_of_Flights[3], f_2016$No_of_Flights[3], f_2017$No_of_Flights[3], f_2018$No_of_Flights[3], f_2019$No_of_Flights[3], f_2020$No_of_Flights[3])

# Total flights averaged across all the months for all the years
y8 = c(mean(f_2015$No_of_Flights), mean(f_2016$No_of_Flights), mean(f_2017$No_of_Flights), mean(f_2018$No_of_Flights), mean(f_2019$No_of_Flights), mean(f_2020$No_of_Flights))


plot(x_axis, y5 , main="Yearly Variation in number of flights", type = "b", col = "red", xlab = "Year", ylab = "Total Flights", ylim=c(2150,2700))
par(new = TRUE)
plot(x_axis, y6 ,type = "b", col = "blue", axes = FALSE, xlab = "Year", ylab = "Total Flights", ylim=c(2150,2700))
par(new = TRUE)
plot(x_axis, y7 ,type = "b", col = "green", axes = FALSE, xlab = "Year", ylab = "Total Flights", ylim=c(2150,2700))
par(new = TRUE)
plot(x_axis, y8 ,type = "b", col = "black", axes = FALSE, xlab = "Year", ylab = "Total Flights", ylim=c(2150,2700))
legend("topleft",legend=c("January", "February", "March", "Average"), col=c("red", "blue", "green", "black"), lty=1:1, cex=0.8)

```


### The future of airlines post COVID-19

```{r,results='markup'}
#Post Covid


library(dplyr)
library(scales)
library(ggplot2)

fancy_scientific <- function(l) {
     # turn in to character string in scientific notation
     l <- format(l, scientific = TRUE)
     # quote the part before the exponent to keep all the digits
     l <- gsub("^(.*)e", "'\\1'e", l)
     # turn the 'e+' into plotmath format
     l <- gsub("e", "%*%10^", l)
     # return this as an expression
     parse(text=l)
}


Monthly_Mean_By_Year = c(mean(p_2015$Total_Passengers), mean(p_2016$Total_Passengers), mean(p_2017$Total_Passengers), mean(p_2018$Total_Passengers), mean(p_2019$Total_Passengers), mean(p_2020$Total_Passengers))

#Monthly_Mean_By_Year
year = c(2015, 2016, 2017, 2018, 2019, 2020)

#plot(year, Monthly_Mean_By_Year, type = 'b')


avg_year<- data.frame(year, Monthly_Mean_By_Year)
avg_year
avg_year %>% 
  ggplot(aes(year, Monthly_Mean_By_Year, colour = year,group=1)) + 
  ggtitle("Plot displaying Average Passengers 2015-2020") +
  geom_line()+
  geom_point()+ scale_y_continuous(labels=fancy_scientific) 
```
```{r,results='markup'}
increase15_16 = (mean(p_2016$Total_Passengers)-mean(p_2015$Total_Passengers))/mean(p_2015$Total_Passengers)*100
increase16_17 = (mean(p_2017$Total_Passengers)-mean(p_2016$Total_Passengers))/mean(p_2016$Total_Passengers)*100
increase17_18 = (mean(p_2018$Total_Passengers)-mean(p_2017$Total_Passengers))/mean(p_2017$Total_Passengers)*100
increase18_19 = (mean(p_2019$Total_Passengers)-mean(p_2018$Total_Passengers))/mean(p_2018$Total_Passengers)*100
increase19_20 = (mean(p_2020$Total_Passengers)-mean(p_2019$Total_Passengers))/mean(p_2019$Total_Passengers)*100

percent_increase = c(increase15_16, increase16_17, increase17_18, increase18_19, increase19_20)
year_range = c(2016, 2017, 2018, 2019, 2020 )
#percent_increase
#plot(year_range, percent_increase, type = 'b')

per <- data.frame(year_range, percent_increase)
per
per %>%
  ggplot(aes(year_range, percent_increase, colour = year_range,group=1)) + 
  ggtitle("Plot displaying percentage change of passengers 2015-2020") +
  geom_line()+
  geom_point()
```
```{r,results='markup'}
avg_flights <- airlines %>%
  group_by(Year) %>%
  summarise(Total_flights = sum(Month))
avg_flights

avg_flights %>% 
  ggplot(aes(Year, Total_flights, colour = Year,group=1)) + 
  ggtitle("Plot displaying Total-Flights 2015-2020") +
  geom_line()+
  geom_point()


yearly_total_flights = c(sum(f_2015$No_of_Flights), sum(f_2016$No_of_Flights), sum(f_2017$No_of_Flights), sum(f_2018$No_of_Flights), sum(f_2019$No_of_Flights), sum(f_2020$No_of_Flights) )


totalFlights_year<- data.frame(year, yearly_total_flights)
totalFlights_year
totalFlights_year %>% 
  ggplot(aes(year, yearly_total_flights, colour = year,group=1)) + 
  ggtitle("Plot displaying yearly total of flights 2015-2020") +
  geom_line()+
  geom_point()+ scale_y_continuous(labels=fancy_scientific) 
```
